A space-time \skewt model for threshold exceedances

### 2016-06-28 LaTeX/skewt_rev.tex
updates to manuscript

\title[A Space-time Skew-$t$ Model for Threshold Exceedances]{A space-time skew-$t$ model for threshold exceedances}
\author
{Samuel A Morris$^{1,*}$\email{samorris@ncsu.edu},
Brian J Reich$^{1}$,
Emeric Thibaud$^{2}$, and
Daniel Cooley$^{2}$\\
$^{1}$Department of Statistics, North Carolina State University, Raleigh, North Carolina, U.S.A. \\
$^{2}$Department of Statistics, Colorado State University, Fort Collins, Colorado, U.S.A.}

\begin{document} %\linenumbers
% \begin{center}
% {\Large {\bf A space-time skew-$t$ model for threshold exceedances}}\\

@@

\begin{abstract}
To assess the compliance of air quality regulations, the Environmental Protection Agency (EPA) must know if a site exceeds a pre-specified level.
In the case of ozone, the level for compliance is fixed at 75 parts per billion, which is high, but not extreme at all locations.
We present a new method based on the spatial skew-$t$ process.
Our method incorporates a random partition to permit long-distance asymptotic independence while allowing for sites that are near one another to be asymptotically dependent, and we incorporate thresholding to allow the tails of the data to speak for themselves.
We also introduce a transformed AR(1) time-series to allow for temporal dependence.
Finally, our model allows for high-dimensional Bayesian inference that is comparable in speed to traditional geostatistical methods for large datasets.
We apply our method to an ozone analysis for July 2005, and find that our model improves over both Gaussian and max-stable methods in terms of predicting exceedances of a high level.
\end{abstract}

\begin{keywords}
Extreme value analysis, MCMC, Random partition, Skew-$t$, Spatio-temporal
\end{keywords}

@@

\section{Introduction}\label{sts:intro}
% In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
Epidemiological studies have linked air quality to public health concerns regarding morbidity and mortality \citep{Samet2000}.
As a result, the Environmental Protection Agency (EPA) has developed a set of standards to help reduce air pollution thereby improving air quality.
Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a high level determined by government regulation.
More specifically, we consider the case of compliance for ozone.
A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over 3 years does not exceed 75 parts per billion (ppb).
\fref{stfig:ozone-10jul} shows the ozone levels from July 10, 2005, at 1089 stations across the United States.
We see a large area above the compliance level in the midwest covering Ohio, Indiana, Illinois, and parts of the surrounding states.
We analyze these data with the goals of spatial prediction for unmonitored locations and to map the probability of extreme events.

@@

A spatial model for threshold exceedances warrants special consideration and standard spatial methods are likely to perform poorly.
First, because we are interested only in high values, we want to ``let the tail speak for itself''.
That is, if we fit a model to the entire data set, low-to-moderate values would influence the fit of the overall model.
As there are more of these values, they can unduly influence the distribution at the higher levels about which we are interested.
Our inference method will only use data which exceed a pre-selected threshold, and will impute data below the threshold, thereby tailoring the fit to the levels of interest.
Second, likelihood-based spatial modeling typically assumes a Gaussian process, which is appropriate when mean behavior is of interest.
However, the Gaussian distribution is light-tailed and symmetric, and therefore may be inappropriate for modeling data which does not share this tail behavior.
Third, we aim to capture the dependence structure when ozone is at high levels, and dependence at these levels may not be well-represented by covariances which focus again on mean behavior.
Asymptotic dependence/independence (see \sref{sts:extdep}) are notions which describe how two random variables' probability of simultaneous exceedance of an extreme threshold behaves as one increases the threshold.
The Gaussian distribution always exhibits asymptotic independence, except in the case of perfect dependence, thus is an inappropriate model for data which exhibits asymptotic dependence.
To allow for more flexibility in the marginal tail and to allow for asymptotic dependence, the skew-$t$ distribution forms the basis for our model.

Our approach differs from threshold modeling approaches based on extreme value distributions.
There has been extensive work on threshold modeling in the field of extreme value statistics where extreme events are naturally defined in terms of exceedances over a high threshold.
\citet{Davison1990} considered modeling threshold exceedances of univariate time series by the generalized Pareto distribution.
Bivariate threshold models for extreme value distributions were considered by \citet{Ledford1996} who introduced a censored approach that provides a way to deal with different types of exceedances of a bivariate threshold in terms of only one or both components.
These threshold models were extended to spatial models for extremes by \citet{Wadsworth2012} and \citet{Thibaud2013} who fit various models to spatial extremes using a censored pairwise likelihood \citep{Padoan2010} based on the approach of \citet{Ledford1996}.
\citet{Huser2014} further extended this to space-time modeling.
\citet{Thibaud2013a}, \citet{Engelke2014}, and \citet{Wadsworth2014}, introduced more efficient inference for threshold exceedances of extremal spatial processes with full likelihood methods.
The previous approaches to threshold modeling are motivated by extreme value theory and assume the threshold is high enough that extremal models are valid for the data, and for extrapolation beyond the range of observed values.
Moreover, these approaches are computationally intensive and limited to rather small datasets.
For example, \citet{Wadsworth2014} present a simulation study with observations at 16 sites on a regular grid, and \citet{Engelke2014} analyze a dataset with observations at 35 meteorological stations.
Our application with ozone data does not fit into this framework because we do not focus on exceedances of a very high threshold, but on exceedances of a fixed threshold.
Furthermore, in our application, we have observations at over 1,000 ozone monitoring locations.

We propose a new spatiotemporal threshold exceedance model based on the skew-$t$ process \citep{Padoan2011}.
\hl{Need to add in more references here to skew-$t$ used in a spatial setting.}
Our model is a threshold exceedance model for the multivariate skew-$t$ distribution using a fixed threshold.
In this setting, we describe the threshold as fixed because it is specified in advance by regulatory compliance.
This differs from the more traditional extremes literature where a threshold is selected to be the value beyond which an extremal model is appropriate for the data.
We use a skew-$t$ distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed level at an unobserved location.

Our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models.
This allows us to use Bayesian methods to impute data below the threshold as well as to make predictions at unobserved locations.
In a spatial setting, the multivariate skew-$t$ distribution demonstrates asymptotic dependence between observations at all sites regardless of the distance between the sites.
In order to address this concern, we introduce a random spatial partition similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
This partitioning alleviates the asymptotic spatial dependence present in the skew-$t$ distribution for sites that are far apart.

The paper is organized as follows.
\sref{sts:spatialskew} is a brief review of the spatial skew-$t$ process.
In \sref{sts:spatial}, we build upon the traditional skew-$t$ process by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
The computing is described in \sref{sts:hier}.
In \sref{sts:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared to Gaussian and max-stable methods.
We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from 800 sites throughout the US in \sref{sts:analysis}.
The final section provides brief discussion and direction for future research.

@@

% In certain settings, the multidimensional distributions are available.
% For example, \citet{Wadsworth2014} discuss a censored Poisson process that exploits the fact that the spectral functions for Brown-Resnick processes are log-Gaussian random fields.
% Additionally, \citep{Engelke2014}, show that upon standardizing marginals to the Gumbel distribution and conditioned on a fixed location exceeding a high threshold, the incremental distribution asymptotically forms a Gaussian process.
% In addition to Brown-Resnick processes, skew elliptical distributions can be used for multivariate modeling with dependent extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
% For example, the skew-normal and skew-$t$ distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate $t$ distributions.
% As with multivariate Gaussian distributions, the multivariate skew-normal distribution demonstrates asymptotic independence. Conversely both the multivariate $t$ and skew-$t$ distributions demonstrate asymptotic dependence \citep{Padoan2011}.
% Additionally, the limiting distribution of the maxima of skew-$t$ random vectors is the extremal skew-$t$ distribution \citep{Padoan2011} of which the extremal-$t$ \citep{Opitz2013} is a special case.

@@

\section{Spatial skew processes}\label{sts:spatialskew}
The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data.
A brief review of the additive process (\citealp{Azzalini2003}; \citealp[p. 129]{Azzalini2014}; \citealp{Beranger2016}) by which a skew-$t$ process is created is given here.

\subsection{Skew-$t$ process} \label{sts:skewt}
Let $Y(\bs)$ be a spatial process defined for spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
The spatial skew-$t$ process can be written
\begin{align}
  Y(\bs) = \bX(\bs)^\top \bbeta + \lambda \sigma |z| + \sigma v(\bs) \label{steq:fullmodel}
\end{align}
$\bX(\bs)$ is the observed covariate vector at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a / 2, b / 2)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, $a$ is the degrees of freedom, $b$ controls the precision of the process, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function $\text{Cor} \left[Y(\bs_1), Y(\bs_2)\right] = \rho(\bs_1, \bs_2)$.
Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
\begin{align}
  \rho(h) = (1 - \gamma) I(h = 0) + \gamma \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right) \label{steq:matern}
\end{align}
where $\varphi$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the Euclidean distance between sites $\bs_1$ and $\bs_2$.

For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^\top$.
After marginalizing over both $z$ and $\sigma$, using the notation from \citet[p. 176]{Azzalini2014},
\begin{align} \label{steq:stazzalini}
  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, a),
\end{align}
that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta \in \calR^n$, variance matrix $\bOmega_{n \times n}$, $\balpha \in \calR^n$ slant parameters, degrees of freedom $a$, where $\bX_{n \times p} = [\bX(\bs_1)^\top, \ldots, \bX(\bs_n)^\top]$,
$\bX(\bs_i), i = 1, \ldots, n$, are observed covariates at each site,
$\bOmega_{n \times n} = \bomega \bar{\bOmega} \bomega$,
$\bar{\bOmega}_{n \times n} = \frac{1}{1 + \lambda^2}(\bSigma + \lambda^2 \bOne \bOne^\top)$,
$\bomega_{n \times n} = \text{diag}\left(\sqrt{\frac{a(1 + \lambda^2)}{ b }}, \ldots, \sqrt{\frac{a(1 + \lambda^2)}{ b }} \right)$,
$\balpha = \lambda (1 + \lambda^2)^{1 / 2} (1 + \lambda^2 \bOne^\top \bSigma^{-1} \bOne)^{-1/2} \bSigma^{-1} \bOne$, and $\bSigma_{n \times n}$ is the positive definite correlation matrix which is obtained from $\rho(h)$.
This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $a$.
Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2014}.
For comparisons with other parameterizations, please see \aref{sta:otherparams}.

@@

\subsection{Extremal dependence}\label{sts:extdep}
Our interest lies in spatial dependence in the tail of the skew-$t$ process.
One measure of extremal dependence is the $\chi$ statistic \citep{Coles1999}.
For a stationary and isotropic spatial process, the $\chi$ statistic for locations $\bs_1$ and $\bs_2$ separated by distance $h = || \bs_1 - \bs_2 ||$ with identical marginal distributions is
\begin{align} \label{steq:chih}
  \chi(h) = \lim_{c \rightarrow c^*} \Pr[Y(\bs_1) > c | Y(\bs_2) > c]
\end{align}
where $c^*$ is the upper limit of the support of $Y$.
For the skew-$t$ distribution, $c^* = \infty$.
If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
For Gaussian processes, $\chi(h) = 0$ regardless of the distance $h$, so they are not suitable for modeling asymptotically dependent extremes.
Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent (the explicit expression for $\chi(h)$ is given in \aref{sta:skewt}).
However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
This occurs because all observations, both near and far, share the same $z$ and $\sigma$ terms.
Therefore, this long-range dependence feature of the skew-$t$ process is not ideal for spatial analysis of large geographic regions where we expect only local spatial dependence.
We propose a solution to this in \sref{sts:part}.

\section{Spatiotemporal skew-$t$ model for threshold exceedances}\label{sts:spatial}
In this section, we propose extensions to the skew-$t$ process to model spatial extremes over a large geographic region by introducing censoring to focus on tail behavior and a random partition to remove long-range asymptotic dependence.
For notational convenience, we introduce the model for a single replication, and then extend this model to the spatiotemporal setting in \sref{sts:temporal}.

\subsection{Censoring to focus on the tails} \label{sts:censoring}
As mentioned previously, we propose to use a censored approach because we are interested in high values and do not want the low-to-moderate values to influence the fit of the overall model.
% The censored observations below the threshold give information on the marginal probabilities to exceed the threshold and on the dependence, but their values are not used to fit the model.
Let
\beq \label{steq:Yt}
  \widetilde{Y}(\bs) = \left\{ \begin{array}{ll}
      Y(\bs) \quad & \delta(\bs) = 1 \\
      T & \delta(\bs) = 0
  \end{array} \right.
\eeq
be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
As our goal is to predict exceedances above a high level $L$, we should select a value for $T \le L$.
For example, in predicting ozone exceedances, we might set $T = 50$ppb in order to predict exceedances of $L = 75$ppb.
Selecting $T$ too small may lead to bias in estimating the tail parameters; selecting $T$ to large increases variance.
We impute the censored values as a step in the MCMC algorithm used to fit the model described in \sref{sts:comp}, and use cross-validation to select $T$.

\subsection{Partitioning to remove long-range asymptotic dependence}\label{sts:part}
The motivation for the partition is that for a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.
Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012} with a hybrid spatial dependence model.
\citet{Huser2014} also allow for varying asymptotic dependence across both space and time with a partition structure represented by random discs moving across the space for a random duration with a random velocity and random radius.
We handle the problem of long-range asymptotic dependence with a random partition.
As discussed in \sref{sts:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site.
The model becomes
\begin{align} \label{steq:partition}
  Y(\bs) &= \bX(\bs)^\top \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
\end{align}
To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu_{PP}$ over spatial domain $\calD \in \calR^2$.
The knots define a random partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
\begin{align} \label{steq:subregions}
  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
\end{align}
In other words, $P_k$ is composed of all sites for which the closest knot is $\bw_k$.
For all $\bs \in P_k$, with $k = 1, 2, \ldots, K$, the functions $z(\bs)$ and $\sigma(\bs)$ are equal to the constants $z_k$ and $\sigma_k$ respectively, and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2_k \iid$ IG$(a / 2, b / 2)$.
So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in \sref{sts:spatialskew}.
Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because $v(\bs)$ spans all partitions.
However, the bivariate distribution for sites in different partitions is neither Gaussian nor skew-$t$ and does not have asymptotic dependence.

>>>>

\title[A Space-time \Skewt Model for Threshold Exceedances]{A space-time \skewt model for threshold exceedances}
\author
{Samuel A Morris$^{1,*}$\email{samorris@ncsu.edu},
Brian J Reich$^{1}$,
Emeric Thibaud$^{2}$, and
Daniel Cooley$^{2}$\\
$^{1}$Department of Statistics, North Carolina State University, Raleigh, North Carolina, U.S.A. \\
$^{2}$Department of Statistics, Colorado State University, Fort Collins, Colorado, U.S.A.}

\begin{document} %\linenumbers
% \begin{center}
% {\Large {\bf A space-time \skewt model for threshold exceedances}}\\

@@

\begin{abstract}
To assess the compliance of air quality regulations, the Environmental Protection Agency (EPA) must know if a site exceeds a pre-specified level.
In the case of ozone, the level for compliance is fixed at 75 parts per billion, which is high, but not extreme at all locations.
We present a new space-time model for threshold exceedances based on the spatial \skewt process.
Our method incorporates a random partition to permit long-distance asymptotic independence while allowing for sites that are near one another to be asymptotically dependent, and we incorporate thresholding to allow the tails of the data to speak for themselves.
We also introduce a transformed AR(1) time-series to allow for temporal dependence.
Finally, our model allows for high-dimensional Bayesian inference that is comparable in speed to traditional geostatistical methods for large datasets.
We apply our method to an ozone analysis for July 2005, and find that our model improves over both Gaussian and max-stable methods in terms of predicting exceedances of a high level.
\end{abstract}

\begin{keywords}
Extreme value analysis, Markov chain Monte Carlo, Random partition, \Skewt, Spatio-temporal modeling
\end{keywords}

@@

\section{Introduction}\label{sts:intro}
% In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
Epidemiological studies have linked air quality to public health concerns regarding morbidity and mortality \citep{Samet2000}.
As a result, the Environmental Protection Agency (EPA) has developed a set of standards to help reduce air pollution thereby improving air quality.
Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a high level determined by government regulation.
More specifically, we consider the case of compliance for ozone.
A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over three years does not exceed 75 parts per billion (ppb).
\fref{stfig:ozone-10jul} shows the ozone levels from July 10, 2005, at 1089 stations across the United States.
We see a large area above the compliance level in the midwest covering Ohio, Indiana, Illinois, and parts of the surrounding states.
We analyze these data with the goals of spatial prediction for unmonitored locations and to map the probability of extreme events.

@@

A spatial model for threshold exceedances warrants special consideration and standard spatial methods are likely to perform poorly.
First, because we are interested only in high values, we want to ``let the tail speak for itself''.
That is, if we fit a model to the entire data set, low-to-moderate values would influence the fit of the overall model.
As there are more of these values, they can unduly influence the distribution at the higher levels about which we are interested.
Our inference method will only use data which exceed a pre-selected threshold and will censor data below the threshold, thereby tailoring the fit to the levels of interest.
Second, likelihood-based spatial modeling typically assumes a Gaussian process, which is appropriate when mean behavior is of interest.
However, the Gaussian distribution is light-tailed and symmetric, and therefore may be inappropriate for modeling data which does not share this tail behavior.
Third, we aim to capture the dependence structure when ozone is at high levels, and dependence at these levels may not be well-represented by covariances which focus again on mean behavior.
Asymptotic dependence/independence (see \sref{sts:extdep}) are notions which describe how two random variables' probability of simultaneous exceedance of an extremely high level.
The Gaussian distribution always exhibits asymptotic independence, except in the case of perfect dependence, thus is an inappropriate model for data which exhibits asymptotic dependence.
To allow for more flexibility in the marginal tail and to allow for asymptotic dependence, the \skewt distribution forms the basis for our model.

Our approach differs from threshold modeling approaches based on extreme value distributions.
There has been extensive work on threshold modeling in the field of extreme value statistics where extreme events are naturally defined in terms of exceedances over a high threshold.
\citet{Davison1990} considered modeling threshold exceedances of univariate time series by the generalized Pareto distribution.
Threshold based inference for multivariate extreme value distributions was considered by \citet{Ledford1996} who introduced a censored approach that provides a way to deal with different types of exceedances of a threshold.
These threshold models were extended to spatial models for extremes by \citet{Wadsworth2012} and \citet{Thibaud2013} who fit various models to spatial extremes using a censored pairwise likelihood \citep{Padoan2010} based on the approach of \citet{Ledford1996}.
\citet{Huser2014} further extended this to space-time modeling.
\citet{Wadsworth2014}, \citet{Engelke2015}, and \citet{Thibaud2013a} introduced more efficient inference for threshold exceedances of extremal spatial processes with full likelihood methods.
The previous approaches to threshold modeling are motivated by extreme value theory and assume the threshold is high enough that extremal models are valid for the data, and for extrapolation beyond the range of observed values.
Moreover, these approaches are computationally intensive and limited to rather small datasets.
For example, \citet{Wadsworth2014} present a simulation study with observations at 16 sites on a regular grid, and \citet{Engelke2015} analyze a dataset with observations at 35 meteorological stations.
Our application with ozone data does not fit into this framework because we do not focus on exceedances of a very high level, and we have observations at over 1,000 ozone monitoring locations.

We propose a new spatiotemporal threshold exceedance model based on the \skewt process \citep{Padoan2011}.
\hl{Need to add in more references here to skew-$t$ used in a spatial setting.}
We use a \skewt distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed level at an unobserved location.
Our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models.
This allows us to use Bayesian methods, which we use to fit the model, handle censored data below the threshold, and make predictions at unobserved locations.
In a spatial setting, the multivariate \skewt distribution demonstrates asymptotic dependence between observations at all sites regardless of the distance between the sites.
In order to address this concern, we introduce a random spatial partition similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
This partitioning alleviates the asymptotic spatial dependence present in the \skewt distribution for sites that are far apart.

The paper is organized as follows.
\sref{sts:spatialskew} is a brief review of the spatial \skewt process.
In \sref{sts:spatial}, we build upon the traditional \skewt process by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
The computing is described in \sref{sts:hier}.
In \sref{sts:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared to Gaussian and max-stable methods.
We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from 1,000 sites throughout the US in \sref{sts:analysis}.
The final section provides brief discussion and direction for future research.

@@

% In certain settings, the multidimensional distributions are available.
% For example, \citet{Wadsworth2014} discuss a censored Poisson process that exploits the fact that the spectral functions for Brown-Resnick processes are log-Gaussian random fields.
% Additionally, \citep{Engelke2014}, show that upon standardizing marginals to the Gumbel distribution and conditioned on a fixed location exceeding a high threshold, the incremental distribution asymptotically forms a Gaussian process.
% In addition to Brown-Resnick processes, skew elliptical distributions can be used for multivariate modeling with dependent extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
% For example, the skew-normal and \skewt distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate $t$ distributions.
% As with multivariate Gaussian distributions, the multivariate skew-normal distribution demonstrates asymptotic independence. Conversely both the multivariate $t$ and \skewt distributions demonstrate asymptotic dependence \citep{Padoan2011}.
% Additionally, the limiting distribution of the maxima of \skewt random vectors is the extremal \skewt distribution \citep{Padoan2011} of which the extremal-$t$ \citep{Opitz2013} is a special case.

@@

\section{Spatial skew processes}\label{sts:spatialskew}
The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data.
A brief review of the additive process (\citealp{Azzalini2003}; \citealp[p. 129]{Azzalini2014}; \citealp{Beranger2016}) by which a \skewt process is created is given here.

\subsection{\Skewt process} \label{sts:skewt}
Let $Y(\bs)$ be a spatial process defined for spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
The spatial \skewt process can be written as
\begin{align}
  Y(\bs) = \bX(\bs)^\top \bbeta + \lambda \sigma |z| + \sigma v(\bs) \label{steq:fullmodel}
\end{align}
where $\bX(\bs)$ is the observed covariate vector at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skewness, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a / 2, b / 2)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, $a$ is the degrees of freedom, $b$ controls the precision of the process, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function $\text{Cor} \left[Y(\bs_1), Y(\bs_2)\right] = \rho(\bs_1, \bs_2)$.
Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
\begin{align}
  \rho(h) = (1 - \gamma) I(h = 0) + \gamma \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right) \label{steq:matern}
\end{align}
where $I(\cdot)$ is an indicator function, $\varphi$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the Euclidean distance between sites $\bs_1$ and $\bs_2$.

For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^\top$, and the covariate matrix $\bX_{n \times p} = [\bX(\bs_1), \ldots, \bX(\bs_n)]^\top$.
After marginalizing over both $z$ and $\sigma$, using the notation from \citet[p. 176]{Azzalini2014},
\begin{align} \label{steq:stazzalini}
  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, a),
\end{align}
that is, $\bY$ follows an $n$-dimensional \skewt distribution with location $\bX \bbeta \in \calR^n$; covariance matrix $\bOmega_{n \times n} = \bomega \bar{\bOmega} \bomega$ where \mbox{$\bar{\bOmega}_{n \times n} = \frac{1}{1 + \lambda^2}(\bSigma + \lambda^2 \bOne \bOne^\top)$}, $\bSigma_{n \times n}$ is the positive definite correlation matrix which is obtained from $\rho(h)$, and $\bomega_{n \times n} = \text{diag}\left(\sqrt{\frac{ b }{a(1 + \lambda^2)}}, \ldots, \sqrt{\frac{ b }{a(1 + \lambda^2)}} \right)$; slant parameters $\balpha \in \calR^n = \lambda (1 + \lambda^2)^{1 / 2} (1 + \lambda^2 \bOne^\top \bSigma^{-1} \bOne)^{-1/2} \bSigma^{-1} \bOne$, and degrees of freedom $a$.
Furthermore, the marginal distributions at each location also follow a univariate \skewt distribution \citep{Azzalini2014}.
This process is desirable because of its flexible tail that is controlled by the skewness parameter and degrees of freedom.
One major difference between the parameterization given in \eref{steq:fullmodel} and other parameterizations is that we characterize the skewnewss using $\lambda$ instead of the slant parameters $\balpha$.
For a more complete comparison with other parameterizations, see \aref{sta:otherparams}.

@@

\subsection{Extremal dependence}\label{sts:extdep}
Our interest lies in spatial dependence in the tail of the \skewt process.
One measure of extremal dependence is the $\chi$ statistic \citep{Coles1999}.
For a stationary and isotropic spatial process, the $\chi$ statistic for locations $\bs_1$ and $\bs_2$ separated by distance $h = || \bs_1 - \bs_2 ||$ with identical marginal distributions is
\begin{align} \label{steq:chih}
  \chi(h) = \lim_{c \rightarrow c^*} \Pr[Y(\bs_1) > c | Y(\bs_2) > c]
\end{align}
where $c^*$ is the upper limit of the support of $Y$.
For the \skewt distribution, $c^* = \infty$.
If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
For Gaussian processes, $\chi(h) = 0$ regardless of the distance $h$, so they are not suitable for modeling asymptotically dependent extremes.
Unlike the Gaussian process, the \skewt process is asymptotically dependent (the explicit expression for $\chi(h)$ is given in \aref{sta:skewt}).
However, one problem with the spatial \skewt process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
This occurs because all observations, both near and far, share the same $z$ and $\sigma$ terms.
Therefore, this long-range dependence feature of the \skewt process is not desirable for spatial analysis of large geographic regions where we expect only local spatial dependence.
We propose a solution to this in \sref{sts:part}.

\section{Spatiotemporal \skewt model for threshold exceedances}\label{sts:spatial}
In this section, we propose extensions to the \skewt process to model spatial extremes over a large geographic region by introducing censoring to focus on tail behavior and a random partition to remove long-range asymptotic dependence.
For notational convenience, we introduce the model for a single replication, and then extend this model to the spatiotemporal setting in \sref{sts:temporal}.

\subsection{Censoring to focus on the tail} \label{sts:censoring}
As mentioned previously, we propose to use a censored approach because we do not want the low-to-moderate values to influence the fit of the overall model.
% The censored observations below the threshold give information on the marginal probabilities to exceed the threshold and on the dependence, but their values are not used to fit the model.
Let
\beq \label{steq:Yt}
  \widetilde{Y}(\bs) = \left\{ \begin{array}{ll}
      Y(\bs) \quad & \delta(\bs) = 1 \\
      T & \delta(\bs) = 0
  \end{array} \right.
\eeq
be the censored process at site $\bs$ where $Y(\bs)$ is the uncensored process, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
As our goal is to model exceedances above a high level $L$, we should select a value for $T \le L$.
For example, in predicting ozone exceedances, we might set $T = 50$ ppb in order to predict exceedances of $L = 75$ ppb.
Selecting $T$ too small may lead to bias in estimating the tail parameters; selecting $T$ too large increases variance.
We impute the censored values as a step in the algorithm used to fit the model described in \sref{sts:comp}, and use cross-validation to select $T$.

\subsection{Partitioning to remove long-range asymptotic dependence}\label{sts:part}
The motivation for the partition is that for a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.
Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012}.
\citet{Huser2014} also allow for varying asymptotic dependence across both space and time with a partition structure represented by random discs moving across the space for a random duration with a random velocity and random radius.
We handle the problem of long-range asymptotic dependence with a random partition.
As discussed in \sref{sts:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site.
The model becomes
\begin{align} \label{steq:partition}
  Y(\bs) &= \bX(\bs)^\top \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
\end{align}
To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process over spatial domain $\calD \in \calR^2$.
The knots define a random partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
\begin{align} \label{steq:subregions}
  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
\end{align}
In other words, $P_k$ is composed of all sites for which the closest knot is $\bw_k$.
For all $\bs \in P_k$, with $k = 1, 2, \ldots, K$, the functions $z(\bs)$ and $\sigma(\bs)$ are equal to the constants $z_k$ and $\sigma_k$ respectively, and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2_k \iid$ IG$(a / 2, b / 2)$.
So, within each partition, $Y(\bs)$ follows the spatial \skewt process defined in \sref{sts:spatialskew}.
Across partitions, the $Y(\bs)$ remain dependence via the correlation function for $v(\bs)$ because $v(\bs)$ spans all partitions.
However, the bivariate distribution for sites in different partitions is neither Gaussian nor \skewt and does not have asymptotic dependence.
